{\LARGE Basis Choice for Scalar-on-Function Regression \\ with Applications to Near-Infrared Spectroscopy}

### 2022-02-07 Paper/main.tex
some correction, interpretation cont.

@@

	In reality, these calculations are often done using basis representations of both the functional principal components $\hat{\nu}^m$ and the observations $x_i(t)$ leading to the following representation. For the sake of clarity, the following equation assumes that the bases used for the expansion of both the observations and the coefficient function are true bases of $\mathbb{L}^2[0,1]$ and can therefore be used to express the corresponding objects exactly.

	\begin{equation}\label{FPCA_basis_expansion}
		\begin{split}
			\hat{\xi}_{i}^m & = \int_{0}^{1} {\color{red}\left( x_i(s) - \hat{\mu}(s)\right)} {\color{blue}\hat{\nu}^m(s)} \mathrm{d}s
			= \int_{0}^{1} {\color{red}\left(\sum_{j \in \mathcal{I}} a_{i,j} \phi_j(s)\right)} {\color{blue}\left(\sum_{k \in \mathcal{L}} b_{k}^m \psi_{k}(s)\right)} \mathrm{d}s \\
			& = \int_{0}^{1} {\color{red}\left(\sum_{j = 1}^{J} a_{i,j} \phi_j(s) + \delta_i^J(s)\right)} {\color{blue}\left(\sum_{k = 1}^{K} b_{k}^m \psi_{k}(s) + \delta_{\beta}^K(s)\right)} \mathrm{d}s \\
			& = \sum_{j = 1}^{J} \left[a_{i,j}\sum_{k = 1}^{K} b_{k}^m \int_{0}^{1} \phi_j(s) \psi_{k}(s)\mathrm{d}s \right] +  \sum_{k = 1}^{K} b_{k}^m \int_{0}^{1} \delta_i^J(s) \psi_{k}(s) \mathrm{d}s + \sum_{j = 1}^{J} a_{i,j} \int_{0}^{1}\phi_j(s) \delta_{\beta}^K(s) \mathrm{d}s
		\end{split}
	\end{equation}

@@

	Truncating the basis used for expansion of the coefficient function already introduces an approximation error.

	\begin{equation}\label{fpcr_reg_both_expansions}
		\begin{split}
			y_i &= \alpha + \int_{0}^{1} \left(\hat{\mu}(s) + \sum_{m = 1}^{N-1} \hat{\xi}^{m}_i \hat{\nu}^{m}(s)\right) {\color{blue}\left(\sum_{j = 1}^{J} b_j \phi_j(s) + \delta_{\beta}^{J}(s)\right)} \mathrm{d}s + \epsilon_i\\
			&= \alpha + \sum_{j = 1}^{J} b_j \int_{0}^{1} \phi_j(s) \hat{\mu}(s) \mathrm{d}s + \int_{0}^{1} \delta_{\beta}^{J}(s) \hat{\mu}(s) \mathrm{d}s + \sum_{m = 1}^{N-1} \left[ \hat{\xi}^m_i \sum_{j = 1}^{J} b_j \int_{0}^{1}\hat{\nu}^m(s) \phi_j(s) \mathrm{d}s \right] \\
			& \quad \quad + \sum_{m = 1}^{N-1} \left[ \hat{\xi}^m_i \int_{0}^{1}\hat{\nu}^m(s) \delta_{\beta}^{J}(s) \mathrm{d}s \right] + \epsilon_i
		\end{split}
	\end{equation}

@@

	\subsection{Generating Similar Curves}
	To avoid small sample problems, we generated 200 similar curves from the gasoline dataset, motivated by \hyperlink{KL}{Karhunen-Lo\'{e}ve Expansion}. First, the initial curves are expressed in terms of a generated cubic B-spline basis which is created using 50 knots. In the R implementation of the fda package that is used, these 50 knots account for 52 basis functions (50+4-2). These smooth curves are then centered, before applying the \hyperlink{KL}{Karhunen-Lo\'{e}ve Expansion}. It is assumed that the scores follow a normal distribution, so the new realizations for the scores are drawn from a multivariate normal distribution $\mathring{\xi} = \left(\mathring{\xi}_{1},\: \dots \:, \mathring{\xi}_{M}\right)' \sim \mathcal{N}(0_M, \; diag(\hat{\lambda}_1,\: \dots\:, \hat{\lambda}_M))$. Finally, we obtain the generated curves $NIR_{sim}$

@@

		\begin{equation}
			\begin{split}
				Y_{1,f} & = \langle NIR, f\rangle + Z  \biggl\lbrack\frac{var(\langle NIR, f\rangle)}{0.9} - var(\langle NIR, f\rangle)\biggr\rbrack \\
				Y_{2,f} & = \langle NIR, f\rangle + Z  \biggl\lbrack\frac{var(\langle NIR, f\rangle)}{0.6} - var(\langle NIR, f\rangle)\biggr\rbrack
			\end{split}
		\end{equation}

@@

		\begin{equation}
			\begin{split}
				\hat{\xi}_{i}^{m, \mathcal{H}} &\approx  \int_{0}^{1} {\color{red} \left(X_{i}^{\mathcal{H}}(s) - \hat{\mu}^{\mathcal{T}}(s)\right)} {\color{blue}\hat{\nu}^{m, \mathcal{T}} }\mathrm{d}s) 
			    = \int_{0}^{1} {\color{red}\left(\sum_{j = 1}^{L} a_{i,j}^{\mathcal{H}} \phi_j(s)\right)} {\color{blue}\left(\sum_{k = 1}^{L} b_{k}^{m, \mathcal{T}} \phi_{k}(s)\right)} \mathrm{d}s \\
				 &\approx \sum_{j = 1}^{L} \left[ a_{i,j}^{\mathcal{H}}\sum_{k = 1}^{L}  b_{k}^{m, \mathcal{T}} \int_{0}^{1} \phi_j(s) \psi_{k}(s)\mathrm{d}s\right]
			\end{split}
		\end{equation}
		with truncation parameter $m$.
		\vspace{0.2cm}

@@

	\subsection{Results}	
	The discussed results and figures of $\hat{\beta}(t)$ for the simulation can be found in the appendix. 
	\subsubsection{Basis Expansion Regression}
	The following results origin from the \hyperref[basis_exp_transf]{Estimation using Basis-Representation}, in which we transform the observed functions to perform regression of a scalar on a countable sequence of regressors, which is then tractable with typical multivariate regression theory.

@@

	A possible explanation applicable for all the specifications might be the effect of the Bias-Variance tradeoff and the following hypothesis: The function $f_1$ is rather simple. meaning that little bias is introduced when choosing a small number of basis functions. For $f_2$, a higher number of basis functions is needed, resulting from the inherent peculiarities of $f_2$. This results in higher numbers of basis functions since the amount of bias is decreasing faster in the number of basis functions than the variance is increasing compared to $f_1$.

	\subsubsection{Functional Principal Component Regression}
	The model which is used for the FPCR is described in \hyperref[fpc_exp_transf]{Estimation using Functional Principal Components}. Additionally to the selection of the truncation parameter $K$, the choice of the number of FPC's adds to the complexity of the model since the eigenvalues and -functions from the decomposition are influenced by the choice of $K$ that was used to approximate the function. But since the FPCR is ultimately estimated with a linear model with the FPC's as regressors, the degree of freedom of these models is not affected by $K$, but only by the number of FPC's $n_{FPC} \in \{2, 3, 4 \}$.

@@

	\subsubsection{Interpretation and Relevance for Application}

	Adress problem of monomials again --> (This could be solved using orthogonal polynomials like Legendre polynomials, see \cite{Dattoli_2001})

	Adress numerical instability?

	maybee both in paragraph under results instead of here?

@@

	\subsection{Interpretation of Results}
		\subsubsection{Basis Expansion Regression}
		\subsubsection{Functional Principal Component Regression}

@@

	\paragraph{Comparison to Penalty Based Smoothing Procedures}

	\paragraph{Larger Range for the Number of Basis Functions for FPCR}

	\paragraph{Larger Range for the Number of Principal Components}

	\paragraph{Input from Physics Analysis}

>>>>

\setlength{\abovedisplayskip}{3pt}
\setlength{\belowdisplayskip}{7pt}

\setlength{\abovedisplayshortskip}{3pt}
\setlength{\belowdisplayshortskip}{7pt}

@@

	In reality, these calculations are often done using basis representations of both the functional principal components $\hat{\nu}^m$ and the observations $x_i(t)$ leading to the following representation. For the sake of clarity, the following equation assumes that the bases used for the expansion of both the observations and the coefficient function are true bases of $\mathbb{L}^2[0,1]$ and can therefore be used to express the corresponding objects exactly.
	\vspace{-0.1cm}
	\begin{equation}\label{FPCA_basis_expansion}
		\begin{split}
			\hat{\xi}_{i}^m & = \int_{0}^{1} {\color{red}\left( x_i(s) - \hat{\mu}(s)\right)} {\color{blue}\hat{\nu}^m(s)} \mathrm{d}s
			= \int_{0}^{1} {\color{red}\left(\sum_{j \in \mathcal{I}} a_{i,j} \phi_j(s)\right)} {\color{blue}\left(\sum_{k \in \mathcal{L}} b_{k}^m \psi_{k}(s)\right)} \mathrm{d}s \\
			& = \int_{0}^{1} {\color{red}\left(\sum_{j = 1}^{J} a_{i,j} \phi_j(s) + \delta_i^J(s)\right)} {\color{blue}\left(\sum_{k = 1}^{K} b_{k}^m \psi_{k}(s) + \delta_{\beta}^K(s)\right)} \mathrm{d}s \\
			& = \sum_{j = 1}^{J} \left[a_{i,j}\sum_{k = 1}^{K} b_{k}^m \int_{0}^{1} \phi_j(s) \psi_{k}(s)\mathrm{d}s \right] +  \sum_{k = 1}^{K} b_{k}^m \int_{0}^{1} \delta_i^J(s) \psi_{k}(s) \mathrm{d}s + \sum_{j = 1}^{J} a_{i,j} \int_{0}^{1}\phi_j(s) \delta_{\beta}^K(s) \mathrm{d}s
		\end{split}
	\end{equation}

@@

	Truncating the basis used for expansion of the coefficient function already introduces an approximation error.
	\vspace{-0.1cm}
	\begin{equation}\label{fpcr_reg_both_expansions}
		\begin{split}
			y_i &= \alpha + \int_{0}^{1} \left(\hat{\mu}(s) + \sum_{m = 1}^{N-1} \hat{\xi}^{m}_i \hat{\nu}^{m}(s)\right) {\color{blue}\left(\sum_{j = 1}^{J} b_j \phi_j(s) + \delta_{\beta}^{J}(s)\right)} \mathrm{d}s + \epsilon_i\\
			&= \alpha + \sum_{j = 1}^{J} b_j \int_{0}^{1} \phi_j(s) \hat{\mu}(s) \mathrm{d}s + \int_{0}^{1} \delta_{\beta}^{J}(s) \hat{\mu}(s) \mathrm{d}s + \sum_{m = 1}^{N-1} \left[ \hat{\xi}^m_i \sum_{j = 1}^{J} b_j \int_{0}^{1}\hat{\nu}^m(s) \phi_j(s) \mathrm{d}s \right] \\
			& \quad \quad + \sum_{m = 1}^{N-1} \left[ \hat{\xi}^m_i \int_{0}^{1}\hat{\nu}^m(s) \delta_{\beta}^{J}(s) \mathrm{d}s \right] + \epsilon_i
		\end{split}
	\end{equation}

@@

	\subsection{Generating Similar Curves}
	To avoid small sample problems, we generated 200 similar curves, $NIR_{sim}$, from the spectra of the gasoline dataset, $NIR$, motivated by \hyperlink{KL}{Karhunen-Lo\'{e}ve Expansion}. First, the initial curves are expressed in terms of a generated cubic B-spline basis which is created using 50 knots. In the R implementation of the fda package that is used, these 50 knots account for 52 basis functions (50+4-2). These smooth curves are then centered, before applying the \hyperlink{KL}{Karhunen-Lo\'{e}ve Expansion}. It is assumed that the scores follow a normal distribution, so the new realizations for the scores are drawn from a multivariate normal distribution $\mathring{\xi} = \left(\mathring{\xi}_{1},\: \dots \:, \mathring{\xi}_{M}\right)' \sim \mathcal{N}(0_M, \; diag(\hat{\lambda}_1,\: \dots\:, \hat{\lambda}_M))$. Finally, we obtain the generated curves $NIR_{sim}$

@@

		\begin{equation}
			\begin{split}
				Y_{1,f} & = \langle NIR, f\rangle + Z  \biggl\lbrack\frac{var(\langle NIR_{sim}, f\rangle)}{0.9} - var(\langle NIR_{sim}, f\rangle)\biggr\rbrack \\
				Y_{2,f} & = \langle NIR_{sim}, f\rangle + Z  \biggl\lbrack\frac{var(\langle NIR_{sim}, f\rangle)}{0.6} - var(\langle NIR_{sim}, f\rangle)\biggr\rbrack
			\end{split}
		\end{equation}

@@

		\begin{equation}
			\begin{split}
				\hat{\xi}_{i}^{m, \mathcal{H},L} &\approx  \int_{0}^{1} {\color{red} \left(X_{i}^{\mathcal{H}}(s) - \hat{\mu}^{\mathcal{T}}(s)\right)} {\color{blue}\hat{\nu}^{m, \mathcal{T}} }\mathrm{d}s) 
			    = \int_{0}^{1} {\color{red}\left(\sum_{j = 1}^{L} a_{i,j}^{\mathcal{H}} \phi_j(s)\right)} {\color{blue}\left(\sum_{k = 1}^{L} b_{k}^{m, \mathcal{T}} \phi_{k}(s)\right)} \mathrm{d}s \\
				 &\approx \sum_{j = 1}^{L} \left[ a_{i,j}^{\mathcal{H}}\sum_{k = 1}^{L}  b_{k}^{m, \mathcal{T}} \int_{0}^{1} \phi_j(s) \psi_{k}(s)\mathrm{d}s\right]
			\end{split}
		\end{equation}

		with truncation parameter $L$.
		\vspace{0.2cm}

@@

	\subsection{Results}	
	The discussed results and figures of $\hat{\beta}(t)$ for the simulation can be found in the appendix. 
	\subsubsection{Basis Expansion Regression}
	The following results origin from the \hyperref[basis_exp_transf]{Estimation using Basis-Representation}, in which we transform the observed functions to perform regression of a scalar on a countable sequence of regressors, which is then tractable with typical multivariate regression theory. examining the results, it appears that the cross-validated MSPE inherets a convex nature in the increasing basis functions.

@@

	\subsubsection{Functional Principal Component Regression}
	The model which is used for the FPCR is described in \hyperref[fpc_exp_transf]{Estimation using Functional Principal Components}. Additionally to the selection of the truncation parameter $K$, the choice of the number of FPC's adds to the complexity of the model since the eigenvalues and -functions from the decomposition are influenced by the choice of $K$ that was used to approximate the function. But since the FPCR is ultimately estimated with a linear model with the FPC's as regressors, the degree of freedom of these models is not affected by $K$, but only by the number of FPC's $n_{FPC} \in \{2, 3, 4 \}$. As will be described, it seems that neither a convex behaviour of the MSPE, nor a clear relationship can be observed between the increasing number of basis function and the number of FPC's.

@@

	\subsubsection{Interpretation and Relevance for Application}
	A possible explanation applicable for all setups using Basis Expansion Regression might be the effect of the Bias-Variance tradeoff and the following hypothesis: For the function $f_1$ it might be that little bias is introduced when choosing a small number of basis functions. For $f_2$, a higher number of basis functions is needed, resulting from the inherent peculiarities of $f_2$. This results in higher numbers of basis functions since the amount of bias is decreasing faster in the number of basis functions than the variance is increasing compared to $f_1$. The uncovered and described convex behaviour of the MSPE also might partly attributed by the bias-variance tradeoff. This could not be observed for the FPCR where no clear relationship between the basis functions and the number of FPC was observed. An hypothesis for this would be that the results for the FPCR are mostly driven by the selected FPC´s. To examine this hypothesis, additional simulations were conducted with a large truncation parameter $L$ for $n_{FPC} \in \{2, 3, 4 \}$, which seem to support this claim (LINK TO APPENDIX).

	%Keeping \cite{Jolliffe_1982} in mind, we ....(refer to problematic of missing insights in pysics / chemistry?)

	%Adress problem of monomials again --> (This could be solved using orthogonal polynomials like Legendre polynomials, see \cite{Dattoli_2001})

	%Adress numerical instability?

	%maybee both in paragraph under results instead of here?

@@

	\subsection{Interpretation of Results}
		\subsubsection{Basis Expansion Regression}
		\subsubsection{Functional Principal Component Regression}
		As for the simulation study, the

@@

	\paragraph{Comparison to Penalty Based Smoothing Procedures}
	\vspace{-0.2cm}

	\paragraph{Larger Range for the Number of Basis Functions for FPCR}
	\vspace{-0.2cm}

	\paragraph{Larger Range for the Number of Principal Components}
	\vspace{-0.2cm}

	\paragraph{Input from Physics Analysis}
	\vspace{-0.2cm}
